An asymptotically normal out-of-sample test of equal predictive accuracy for nested models

### 2012-03-30 Paper.tex
Consolidated and cleaned up source files
- Added DELETE_ON_ERROR target to makefile
- switched versioning to date-based and put in its own file
Added material for multiple tests and for empirics

\author{Gray
  Calhoun\thanks{\texttt{http://www.econ.iastate.edu/$\sim$gcalhoun};
    Economics Department; Iowa State University; Ames, IA 50011.
    Telephone: (515) 294-6271.  Email: \texttt{gcalhoun@iastate.edu}.}
  \\ Iowa State University}

@@

\begin{abstract} 
  \noindent This paper proposes a modification of \citepos[\textit{J.
      Econom.}]{ClW:07} adjusted out-of-sample $t$-test.  The
  alternative model is still estimated with a fixed-length rolling
  window, but the benchmark is estimated with a recursive window. The
  resulting statistic is asymptotically normal even when the models
  are nested.  Moreover, the alternative model can be estimated using
  popular model selection methods, such as the \aic\ or \bic.  This
  paper also presents a new bootstrap procedure based on
  \citet[\textit{Econometrica}]{RoW:05} to compare multiple models
  while controlling family-wise error that substantially improves
  existing block bootstrap methods for out-of-sample statistics.  This
  procedure is then used to analyze
  \citepos[\textit{Rev. Finan. Stud.}]{GoW:08} excess returns dataset.

@@

\noindent Keywords: Forecast Evaluation, Martingale Difference
Sequence, Model Selection, Family-Wise Error Rate; Multiple Testing;
Bootstrap; Reality Check

@@

\newpage \noindent In a pair of papers, \citet{ClW:06,ClW:07} develop
an out-of-sample (\oos) test of the null hypothesis that a small
benchmark model is correctly specified.  Their test compares the
forecasting performance of a pair of nested models, and the null
hypothesis is that the innovations in the smaller model form a
martingale difference sequence.  This test procedure is
popular,\footnote{As of 03 June 2011, \citet{ClW:06} has been cited by
  26 papers indexed by the \textsc{isi} Web of Knowledge, and
  \citet{ClW:07} by 37. They have been cited by 113 and 165 papers
  respectively indexed by Google Scholar.} and one assumes that this
is due in part to the statistic's convenience.  The statistic is
approximately normal after adjusting for the estimation error of the
larger model.  Normality comes from a fixed-length rolling window, as
in \citet{GiW:06}, and the adjustment centers the statistic
appropriately.  This statistic is especially convenient because other
tests for similar hypotheses \citep[among
  others]{Mcc:07,ClM:05,ClM:01,CoS:04,CoS:02,CCS:01} have a
nonstandard limit distribution and place restrictions on the models
under consideration, while other asymptotically normal statistics test
a different null hypothesis \citep{GiW:06} or place assumptions on the
models and \dgp\ that are rarely appropriate in empirical work
\citep{Wes:96,WeM:98,Mcc:00,DiM:95}.

@@

In this paper, I show that a modified version of their statistic is
asymptotically normal even when the smaller model is estimated.  To
achieve normality, we need a consistent estimate of the pseudo-true
benchmark model, while maintaining an inconsistent estimate of the
larger model so that we can ignore nesting.  We can meet both needs by
using different window strategies for each model; the benchmark model
is estimated using a recursive window and the alternative with a
fixed-length rolling window.

@@

Statistics for comparing two models are useful, but a researcher often
has a set of potential models, and wants to know which of them
significantly outperform the benchmark.  As \citet{Whi:00} shows,
looking at the naive $p$-values is misleading, but bootstrapping the
statistics can lead to a valid test (White calls this procedure the
``bootstrap reality check'').  So I present a test for equal
predictive ability of multiple models based on \citepos{RoW:05} StepM,
which uses a step-down procedure that iteratively rejects models to
achieve higher power than the reality check (see Theorem~\ref{res:2}
for details) and will indicate which of the models improves on the
benchmark.

I also present a new result for the validity of using the stationary
boostrap with \oos\ statistics, which is necessary to verify that the
StepM is appropriate.  Existing results on boostrapping
\oos\ statistics have some drawbacks.  \citet{Whi:00} and
\citet{Han:05} use the stationary bootstrap, but require the test
sample to be much smaller than the training sample, which is obviously
inappropriate here.  \citet{CoS:07} relax that requirement, but only
consider the moving block bootstrap \citep{LiS:92,Kun:89} and make the
statistic more complicated than necessary by adjusting the objective
function of the bootstrapped statistic to center it
correctly.\footnote{Their recentering is required for consistency and
  does not give higher order accuracy.}  A parametric bootstrap, such
as that used by \citet{ClM:11}, would be difficult to implement here
since we want to test whether several different null hypotheses hold
individually (as I'll discuss later, Clark and McCracken test a single
null hypothesis that describes the behavior of all of the models under
consideration and does not let the researcher know which of the models
outperforms, only that at least one does).  So this paper's result for
the stationary bootstrap of \oos\ statistics considerably improves on
existing procedures.

@@

\begin{thm}\label{res:1}
  Suppose that we have two models $\hat{y}_{0t}$ and $\hat{y}_{1t}$ to
  forecast the variable $y_t$, and have observations for
  $t=1,\dots,T+1$.  Assume the following hold:
  \begin{enumerate}
  \item \label{item:1} The benchmark forecast $\hat{y}_{0,t+1}$, is
    estimated using \ols\ with a recursive window: $\hat{y}_{0,t+1} =
    x_t'\hat{\beta}_t$ for some vector of predictors $x_t$ with
    \begin{equation}
      \hat{\beta}_t = \Big(\sum_{s=2}^{t} x_{t-1} x_{t-1}'\Big)^{-1}
      \sum_{s=2}^t x_{t-1} y_t, \qquad t = R,\dots,T.
    \end{equation}
    Also define $\beta_0 = \big(\sum_{t=2}^{T} \E x_{t-1}
    x_{t-1}'\big)^{-1} \sum_{t=2}^T \E x_{t-1} y_t$.
  \item \label{item:2} The alternative forecast, $\hat{y}_{1t}$, is
    estimated using a rolling window of fixed length $R$ (which is
    less than $T$), so $\hat{y}_{1,t+1} =
    \psi(y_t,z_t,\dots,y_{t-R+1}, z_{t-R+1})$ where $\psi$ is a known
    function and $z_t$ is a sequence of predictors that includes
    $x_t$.
  \item \label{item:3} The series $y_t$ and $z_t$ are \ned\ of size
    $-1/2$ on a strong mixing series of size $-r/(r-2)$ for $r>2$ or a
    uniform mixing series of size $-r/2(r-1)$.  Also, $y_t$,
    $\hat{y}_{1t}$, and $x_t$ have uniformly bounded $2 r+\delta$
    moments for some positive $\delta$.
  \item \label{item:4} Define \[f_t(\beta) = (y_{t+1} - x_t'\beta)^2 -
    (y_{t+1} - \hat{y}_{1,t+1})^2 + (x_t'\beta - \hat{y}_{1,t+1})^2,\]
    $\bar{f}(\beta) = P^{-1} \sum_{t=R+1}^T f_t(\beta)$, and $\bar f =
    P^{-1} \sum_{t=R+1}^{T} f_t(\hat{\beta}_t)$, with $P = T - R$. The
    processes $f_t(\beta_0)$ and $x_t(y_{t+1} - x_t'\beta_0)$ are
    covariance stationary and $\bar f(\beta_0)$ has uniformly positive
    long run variance.
  \end{enumerate}
  Under the null hypothesis that $y_t - \hat{y}_{0t}(\beta_0)$ is a
  martingale difference sequence with respect to the filtration
  $\mathcal{F}_t = \sigma(y_t, z_{t-1}, y_{t-1}, z_{t-2},\dots)$, $\sqrt{P} \bar f /
  \hat{\sigma} \to^d N(0,1)$, where
  \begin{align*}
    \hat{\sigma}^2 &= \hat{S}_{ff} + 2 \Pi (\hat{S}_{fg} + \hat{S}_{gg}), &
    \hat{S}_{ff} &= \frac1P \sum_{t=R+1}^T (f_t(\hat{\beta}_t) - \bar
    f)^2, \\
    \hat{S}_{fg} &= \frac1P \sum_{t=R+1}^T (f_t(\hat{\beta}_t) -
    \bar{f})(g_t(\hat{\beta}_t) - \bar g)', &
    \hat{S}_{gg} &= \frac1P \sum_{t=R+1}^T (g_t(\hat{\beta}_t) - \bar
    g)(g_t(\hat{\beta}_t) - \bar g)',
  \end{align*}
  $\Pi = 1 - (R/P) \log(1 + P/R)$, $\bar{g} = T^{-1}
  \sum_{t=R+1}^T g_t(\hat{\beta}_t)$, $\X' = [x_1,\dots,x_T]$, and
  \begin{equation*}
  g_t(\beta) = \Bigg\{\frac{2}{P}\sum_{t=R}^{T-1} x_t (2 x_t'\beta -
    y_{t+1} - \hat{y}_{1,t+1}) \Bigg\} \big(\X'\X / T \big)^{-1}  x_t(y_{t+1} - x_t'\beta).
  \end{equation*}
\end{thm}

@@

I also use Lemma~\ref{res:1} to test multiple models at once.  There
are two approaches one can take.  The first is to test whether any of
the models can improve on the benchmark; \citet{Whi:00},
\citet{Han:05}, and \citet{ClM:11} take this approach.  This approach
requires testing a single null hypothesis, that the best model is no
better than the benchmark, and one wants a test statistic to control
size as usual.  The second approach, taken by \citet{RoW:05}, is to
try to determine which models outperform the benchmark, and to find as
many as possible that do.  Now there are as many hypotheses as there
are models, so we need to control a quantity similar to size, the
family-wise error rate, which is the probability that one or more true
hypotheses is rejected.  To appreciate the difference between these
two approaches, suppose that there are two models and the first
outperforms a benchmark but the second doesn't.  Under the first
approach, finding that both models outperform the benchmark is fine:
the null (that none of them do) is false so we want to reject.  But
under the second approach, this finding is a failure, as we
incorrectly believe that the second model beats the benchmark.

While methods like \citepos{Whi:00} and \citepos{Han:05} control the
family-wise error, \citepos{RoW:05} StepM procedure will typically find
more outperforming models while still controlling the family-wise error
rate.  All of these methods rely on the bootstrap to estimate the
joint dependence between the test statistic associated with each
model.  The bootstrap we will use in this paper is new, and is
outlined in Algorithm~\ref{alg:1}.
\begin{alg}\label{alg:1}
  \begin{enumerate}
  \item Define the data matrices $\X_\R = (X_1,\dots,X_R)'$ and $\X_{\p} =
    (X_{R+1},\dots,X_{T-1})$ with
    \begin{equation*}
      X_t = \begin{cases}
        (y_{t+1}, x_t')' & t \leq R \\
        (y_{t+1}, x_t', \hat{y}_{1,t+1})' & t > R.
      \end{cases}
    \end{equation*}
  \item Draw $B$ samples of $P$ observations from $\X_{\p}$ using the
    stationary bootstrap with expected block size $b$ and denote each
    sample as $\X_{\p, l}^{*}$; let $\X_l^{*} = [\X_{\R}', \X_{\p,l}^{*\prime}]'$.
  \item Estimate $\bar{f}^{*}_l$ and $\hat{\sigma}_l^{*}$ as in
    Theorem~\ref{res:1} for each bootstrap sample $\X^{*}_l$.
  \end{enumerate}
\end{alg}

Theorem~\ref{res:2} details the final procedure.

\begin{thm}\label{res:2}
  Suppose the conditions of Lemma~\ref{res:1} hold, with assumptions
  on the alternative model, $\hat{y}_{1t}$, holding for each of the
  $m$ models $\hat{y}_{1t}, \dots, \hat{y}_{mt}$, let the subscript
  $l$ denote the different quantities associated with $\hat{y}_{lt}$
  (i.e. $\mathcal{F}_{lt}$, etc.), and let $H_l$ be the null
  hypothesis that $y_t - \hat{y}_{0t}$ is an \mds\ with respect to
  $F_{lt}$.  Also assume that the long-run variance-covariance matrix
  of the vector $(\bar{f}_1,\dots,\bar{f}_m)'$ is uniformly positive
  definite.

  Let the $*$-superscript denote a random variable with the
  distribution induced by the stationary bootstrap of
  Algorithm~\ref{alg:1}, letting $X_t = (y_{t+1}, x_t', \hat{y}_{1,t+1},
  \dots, \hat{y}_{m, t+1})'$ for $t > R$ (and unchanged for $t \leq
  R$).  Define $M_j$ as follows:
  \begin{enumerate}
  \item define $M_0 = \emptyset$,
  \item for $j = 1,2,\dots,m$, let $M_j = \{i : \bar{f}_{i} /
    \hat{\sigma}_i > \hat{d}_j\}$, where $\hat{d}_j$ is the $1-\alpha$
    quantile of $\max_{i \notin M_{j-1}} (\bar{f}_{i}^{*} -
    \bar{f_i}(\hat{\beta}_{T+1})) / \hat{\sigma}_i^{*}$,
  \end{enumerate}
  and reject all of the hypotheses $H_i$ with $i \in \bigcup_{j=0}^m
  M_j$.

  If $b \to \infty$ and $b/P \to 0$, this algorithm controls the
  \fwe\ at level $\alpha$ for the family of null hypotheses $H_i$
  ($i=1,\dots,m$): the probability of rejecting any true $H_i$ is less
  than or equal to $\alpha$.  Moreover, the probability that $H_i$ is
  rejected converges to one if $y_{t} - \hat{y}_{0t}(\beta_0)$ and
  $\hat{y}_{it}$ have nonzero correlation.
\end{thm}

@@

\section{General Results for the Bootstrap}\label{sec:1b}
The validity of the stationary boostrap in Theorem~\ref{res:2} is a
special case of a result of independent interest---the validity in
general of the stationary bootstrap when \oos\ statistics are
asymptotically normal.  Existing results---\citet{Whi:00},
\citet{Han:05}, and \citet{CoS:07}---are not appropriate here;
\citet{Whi:00} and \citet{Han:05} require that $R$ be very large,
while this paper requires $R$ to be finite; and \citet{CoS:07} require
a complicated modification of the statistic of interest be used to
create the bootstrap distribution.  Instead of using these results,
I'll present one more general.  The proof largely follows the approach
of \citet{Wes:96}, \citet{WeM:98}, and \citet{Mcc:00}, using the
arguments of \citet{Cal:11d} for asymptotic normality.

In this section, let $\to^{p^{*}}$ and $\to^{d^{*}}$ refer to
convergence in probability or distribution conditional on the observed
data.  Similarly, $\E^{*}$, $\var^{*}$, and $\cov^{*}$ refer to the
expectation, variance, and covariance with respect to the probability
measure induced by the stationary bootstrap, and $y_t^{*}$, etc. is
the random variable $y_t$ but under the bootstrap-induced \cdf.

Note that the notation in this section is more general than that of
Section~\ref{sec:1}.  The assumptions required are generalizations of
the conditions of Theorems~\ref{res:1} and~\ref{res:2}.  See
\citet{Wes:96,Wes:06}, \citet{WeM:98}, and \citet{Mcc:00} for a
discussion of these conditions, as theirs are nearly identical.

\begin{asmp}\label{asmp:a1} 
  Let $\{y_t, z_t\}$ be an $L_2$-\ned\ process of size $-1/2$ on a
  strong mixing series $\{V_t\}$ of size $-r/(r-2)$, with $r > 2$.
\end{asmp}

@@

\begin{thm}\label{res:3}
  If Assumptions~\ref{asmp:a1}--\ref{asmp:a5} hold , $P, b \to \infty$
  and $b/P \to 0$ as $T \to \infty$, and (p.d. assumption holds) then
  \begin{equation}
    \pr[\sup_x | \pr^*[\sqrt{P} (\bar{f}^* - \bar{f}(\hat{\beta}_{T+1}))
        \leq x] - \pr[\sqrt{P}( \bar{f} - \E f(\beta_0)) \leq x] | >
      \epsilon] \to 0
  \end{equation}
  for all $\epsilon > 0$.
\end{thm}

@@

\begin{lema}\label{res:a5}
  If Assumptions~\ref{asmp:a1}, \ref{asmp:a3}, \ref{asmp:a4}, and
  \ref{asmp:a5} hold, and instead of \ref{asmp:a2} we have the weaker
  condition: The estimator $\hat{\beta}_t$ of $\beta_0$ satisfies
  $\hat{\beta}_{t} - \beta_{0} = \hat{B}_{t} H_t$; $\hat{B}_{t}$ is a
  sequence of $k$ by $q$ matrices that converges almost surely to the
  deterministic sequence $B_{t}$, in the sense that $\sup_{t \in
    \{S_T,\dots,T\}} | \hat{B}_{t} - B | \to 0$ almost surely for any
  sequence $S_T$ that diverges to $\infty$ with $T$; $H_{t}$ is a
  sequence of $q$-vectors such that $H_{t} = (1/t) \sum_{s=1}^t
  h_{s}(\beta_{0})$, $h_{s}(\beta) = h(y_{s}, z_{s}, \beta)$, and $\E
  h_{s}(\beta_{0}) = 0$ for all $s$.  And $P \to \infty$ as $T \to
  \infty$, (add p.d. assumption too) then
  \begin{equation}
    P^{-1/2} \sum_{t=R+1}^T
    (f_{t}(\hat{\beta}_{t}) - \E f_{t}(\beta_{0})) \to^d N(0, \Omega),
  \end{equation}
  with
  \begin{align}
    \Omega =     
  \end{align}
\end{lema}
\begin{proof}[Proof of Lemma~\ref{res:a5}]
  Let $R'$ satisfy $R' \to \infty$ as $T \to \infty$ and $R'/T \to 0$
  and define $P' = T - R'$.  Then we can replace $P$ and $R$ with $P'$
  and $R'$ in the preceding equations without changing the asymptotic
  distributions of the statistics.  Note that $P'/R' \to \infty$.  The
  proof then follows exactly as in \citet[Theorem 2.3.1]{Mcc:00}
\end{proof}

>>>>

\author{Gray Calhoun\thanks{Address: 467 Heady Hall, Economics
    Department; Iowa State University; Ames, IA 50011. Telephone:
    (515) 294-6271.  Email: \texttt{gcalhoun@iastate.edu}.} \\ Iowa
  State University}

@@

\begin{abstract} 
  \noindent This paper proposes a modification of \citepos[J.
  Econom.]{ClW:07} adjusted out-of-sample $t$-test.  The
  alternative model is still estimated with a fixed-length rolling
  window, but the benchmark is estimated with a recursive window. The
  resulting statistic is asymptotically normal even when the models
  are nested.  Moreover, the alternative model can be estimated using
  popular model selection methods, such as the \aic\ or \bic.  This
  paper also presents a new bootstrap procedure based on
  \citet[Econometrica]{RoW:05} to compare multiple models while
  controlling family-wise error, and demonstrates this procedure by
  analyzing \citepos[Rev. Finan. Stud.]{GoW:08} excess returns dataset.

@@

\noindent Keywords: Forecast Evaluation, Martingale Difference
Sequence, Model Selection, Family-wise Error; Multiple Testing;
Bootstrap; Reality Check

@@

\newpage \noindent In a pair of papers, \citet{ClW:06,ClW:07} develop
an out-of-sample (\oos) test of the null hypothesis that a small
benchmark model is correctly specified.  Their test compares the
forecasting performance of a pair of nested models, and the null
hypothesis is that the innovations in the smaller model form a
martingale difference sequence.  This test procedure is
popular,\footnote{As of 03 June 2011, \citet{ClW:06} has been cited by
  26 papers indexed by the \textsc{isi} Web of Knowledge, and
  \citet{ClW:07} by 37. They have been cited by 113 and 165 papers
  respectively indexed by Google Scholar.} and one assumes that this
is due in part to the statistic's convenience.  The statistic is
approximately normal after adjusting for the estimation error of the
larger model.  Normality comes from a fixed-length rolling window, as
in \citet{GiW:06}, and the adjustment centers the statistic at the
appropriate population quantities.  This statistic is especially
convenient because other tests for similar hypotheses \citep[among
others]{Mcc:07,ClM:05,ClM:01,CoS:04,CoS:02,CCS:01} have a nonstandard
limit distribution and place restrictions on the models under
consideration, while other asymptotically normal statistics test a
different null hypothesis \citep{GiW:06} or place assumptions on the
models and \dgp\ that are rarely appropriate in empirical work
\citep{Wes:96,WeM:98,Mcc:00,DiM:95}.

@@

In this paper, I prove that a modified version of their statistic is
asymptotically normal even when the smaller model is estimated.  To
achieve normality, we need a consistent estimate of the pseudo-true
benchmark model, while maintaining an inconsistent estimate of the
larger model so that we can ignore nesting.  We can meet both needs by
using different window strategies for each model; the benchmark model
is estimated using a recursive window and the alternative with a
fixed-length rolling window.

@@

Statistics for comparing two models are useful, but a researcher often
has a set of potential models, and wants to know which of them
significantly outperform the benchmark.  As \citet{Whi:00} shows,
looking at the naive $p$-values is misleading, but bootstrapping the
statistics can give a valid test (White calls this procedure the
``bootstrap reality check'').  So I present a test for equal
predictive ability of multiple models based on \citepos{RoW:05} StepM,
which uses a step-down procedure that iteratively rejects models to
achieve higher power than the reality check (see Theorem~\ref{res:2}
for details) and will indicate which of the models improves on the
benchmark.

I also present a new result for the validity of using the stationary
boostrap with \oos\ statistics, which is necessary to verify that the
StepM is appropriate.  Existing results on boostrapping \oos\
statistics have some drawbacks.  \citet{Whi:00} and \citet{Han:05} use
the stationary bootstrap, but require the test sample to be much
smaller than the training sample, which is obviously inappropriate
here.  \citet{CoS:07} relax that requirement, but only consider the
moving block bootstrap \citep{LiS:92,Kun:89} and make the statistic
more complicated than necessary by adjusting the objective function of
the bootstrapped statistic to center it correctly.\footnote{Their
  recentering is required for consistency and does not give higher
  order accuracy.}  A parametric bootstrap, such as that used by
\citet{ClM:11}, would be difficult to implement here since we want to
test whether several different null hypotheses hold individually (as
I'll discuss later, Clark and McCracken test a single null hypothesis
that describes the behavior of all of the models under consideration).
So this paper's result for the stationary bootstrap of \oos\ statistics
improves on existing procedures.

@@

\begin{thm}\label{res:1}
  Suppose that we have two models $\hat{y}_{0t}$ and $\hat{y}_{1t}$ to
  forecast the variable $y_t$, and have observations for
  $t=1,\dots,T+1$.  Assume the following hold:
  \begin{enumerate}
  \item \label{item:1} The benchmark forecast $\hat{y}_{0,t+1}$, is
    estimated using \ols\ with a recursive window: $\hat{y}_{0,t+1} =
    x_t'\hat{\beta}_t$ for some vector of predictors $x_t$ with
    \begin{equation}
      \hat{\beta}_t = \Big(\sum_{s=2}^{t} x_{t-1} x_{t-1}'\Big)^{-1}
      \sum_{s=2}^t x_{t-1} y_t, \qquad t = R,\dots,T.
    \end{equation}
    Also define $\beta_0 = \big(\sum_{t=2}^{T} \E x_{t-1}
    x_{t-1}'\big)^{-1} \sum_{t=2}^T \E x_{t-1} y_t$.
  \item \label{item:2} The alternative forecast, $\hat{y}_{1t}$, is
    estimated using a rolling window of fixed length $R$ (which is
    less than $T$), so $\hat{y}_{1,t+1} =
    \psi(y_t,z_t,\dots,y_{t-R+1}, z_{t-R+1})$ where $\psi$ is a known
    function and $z_t$ is a sequence of predictors that includes
    $x_t$.
  \item \label{item:3} The series $y_t$ and $z_t$ are \ned\ of size
    $-1/2$ on a strong mixing series of size $-r/(r-2)$ for $r>2$.
    Also, $y_t$, $\hat{y}_{1t}$, and $x_t$ have uniformly bounded $2
    r+\delta$ moments for some positive $\delta$.
  \item \label{item:4} Define \[f_t(\beta) = (y_{t+1} - x_t'\beta)^2 -
    (y_{t+1} - \hat{y}_{1,t+1})^2 + (x_t'\beta - \hat{y}_{1,t+1})^2,\]
    $\bar{f}(\beta) = P^{-1} \sum_{t=R+1}^T f_t(\beta)$, and $\bar f =
    P^{-1} \sum_{t=R+1}^{T} f_t(\hat{\beta}_t)$, with $P = T - R$. The
    processes $f_t(\beta_0)$ and $x_t(y_{t+1} - x_t'\beta)_0$ are
    covariance stationary and $\bar f(\beta_0)$ has uniformly positive
    long run variance.
  \end{enumerate}
  Under the null hypothesis that $y_t - \hat{y}_{0t}(\beta_0)$ is a
  martingale difference sequence with respect to the filtration
  $\mathcal{F}_t = \sigma(y_t, z_{t-1}, y_{t-1}, z_{t-2},\dots)$, $\sqrt{P} \bar f /
  \hat{\sigma} \to^d N(0,1)$, where
  \begin{align*}
    \hat{\sigma}^2 &= \hat{S}_{ff} + 2 \Pi (\hat{S}_{fg} + \hat{S}_{gg}), &
    \hat{S}_{ff} &= \frac1P \sum_{t=R+1}^T (f_t(\hat{\beta}_t) - \bar
    f)^2, \\
    \hat{S}_{fg} &= \frac1P \sum_{t=R+1}^T (f_t(\hat{\beta}_t) -
    \bar{f})(g_t(\hat{\beta}_t) - \bar g)', &
    \hat{S}_{gg} &= \frac1P \sum_{t=R+1}^T (g_t(\hat{\beta}_t) - \bar
    g)(g_t(\hat{\beta}_t) - \bar g)',
  \end{align*}
  $\Pi = 1 - (R/P) \log(1 + P/R)$, $\bar{g} = T^{-1}
  \sum_{t=R+1}^T g_t(\hat{\beta}_t)$, $\X' = [x_1,\dots,x_T]$, and
  \begin{equation*}
  g_t(\beta) = \Bigg\{\frac{2}{P}\sum_{t=R}^{T-1} x_t (2 x_t'\beta -
    y_{t+1} - \hat{y}_{1,t+1}) \Bigg\} \big(\X'\X / T \big)^{-1}  x_t(y_{t+1} - x_t'\beta).
  \end{equation*}
\end{thm}

@@

Before presenting the method for testing multiple models, I'll give a
version of \citepos{PoR:94} stationary bootstrap for the \oos\
statistic of Theorem~\ref{res:1}.  Note that the result holds as well
for the moving blocks bootstrap of \citet{Kun:89} and \citet{LiS:92}.
This result will be used in Theorem~\ref{res:2} to derive the joint
distribution of several models' \oos\ statistics.  The
boostrap-induced probability measure is denoted $\pr^{*}$.

\begin{lem}\label{res:3}
  Suppose the conditions of Theorem~\ref{res:1} hold, but strengthen
  condition~\ref{item:3} so that $y_t$ and $z_t$ are \ned\ of size
  $-1$ on a strong mixing series of size $-(2r + \rho)/(r-2)$ with
  $\rho > 0$ and $r > 2$.  Consider the following procedure:
  \begin{enumerate}
  \item Define the data matrices $\X_\R = (X_1,\dots,X_R)'$ and $\X_{\p} =
    (X_{R+1},\dots,X_{T-1})$ with
    \begin{equation*}
      X_t = \begin{cases}
        (y_{t+1}, x_t')' & t \leq R \\
        (y_{t+1}, x_t', \hat{y}_{1,t+1})' & t > R.
      \end{cases}
    \end{equation*}
  \item Draw $B$ samples of $P$ observations from $\X_{\p}$ using the
    stationary bootstrap with expected block size $b$ and denote each
    sample as $\X_{\p, l}^{*}$; let $\X_l^{*} = [\X_{\R}', \X_{\p,l}^{*\prime}]'$.
  \item Estimate $\bar{f}^{*}_l$ and $\hat{\sigma}_l^{*}$ as in
    Theorem~\ref{res:1} for each bootstrap sample $\X^{*}_l$.
  \end{enumerate}

  Let $(\bar{f}^{*},\hat{\sigma}^{*})$ be random variables with the
  same (conditional) distribution as $(\bar{f}_l^{*},
  \hat{\sigma}^{*}_l)$.  If $b \to \infty$ and $b^2/P \to 0$
  \begin{gather}
    \pr\big[\sup_x \big| \pr^{*}\big[\sqrt{P}(\bar{f}^{*} -
    \bar{f}(\hat{\beta}_{T+1})) \leq x \big] - \pr\big[\sqrt{P}
    (\bar{f} - \E \bar{f}(\beta_0)) \leq x\big] \big| > \epsilon\big]
    \to 0 \intertext{and} \pr\Bigg[\sup_x \Bigg|
    \pr^{*}\Bigg[\frac{\bar{f}^{*} -
      \bar{f}(\hat{\beta}_{T+1})}{\hat{\sigma}^{*}/\sqrt{P}} \leq x
    \Bigg] - \pr\Bigg[\frac{\bar{f} - \E
      \bar{f}(\beta_0)}{\hat{\sigma}/\sqrt{P}} \leq x\Bigg] \Bigg| >
    \epsilon\Bigg] \to 0
  \end{gather}
  for all $\epsilon>0$.
\end{lem}

\begin{rem}
  It is likely that Lemma~\ref{res:3} holds even if $\sqrt{P} \bar{f}$
  is not asymptotically normal, but I leave that case to future
  research as it's not necessary for this paper's results.
\end{rem}

Finally, I use Lemma~\ref{res:3} to test multiple models at once.
There are two approaches one can take.  The first is to test whether
any of the models can improve on the benchmark; \citet{Whi:00},
\citet{Han:05}, and \citet{ClM:11} take this approach.  This approach
requires testing a single null hypothesis, that the best model is no
better than the benchmark, and one wants a test statistic to control
size as usual.  The second approach, taken by \citet{RoW:05}, is to
try to determine which models outperform the benchmark, and to find as
many as possible that do.  Now there are as many hypotheses as there
are models, so we need to control a quantity similar to size, the
family-wise error rate, which is the probability that one or more true
hypotheses is rejected.  To appreciate the difference between these
two approaches, suppose that there are two models and the first
outperforms a benchmark but the second doesn't.  Under the first
approach, finding that both models outperform the benchmark is fine:
the null (that none of them do) is false so we want to reject.  But
under the second approach, this finding is a failure, as we
incorrectly believe that the second model beats the benchmark.

While methods like \citepos{Whi:00} and \citepos{Han:05} control the
family-wise error, \citepos{RoW:05} StepM procedure will typically find
more outperforming models while still controlling the family-wise error
rate.  Theorem~\ref{res:2} shows how the \oos\ bootstrap of
Lemma~\ref{res:3} can be used with the StepM procedure.

\begin{thm}\label{res:2}
  Suppose the conditions of Lemma~\ref{res:3} hold, with assumptions
  on the alternative model, $\hat{y}_{1t}$, holding for each of the
  $m$ models $\hat{y}_{1t}, \dots, \hat{y}_{mt}$, let the subscript
  $l$ denote the different quantities associated with $\hat{y}_{lt}$
  (i.e. $\mathcal{F}_{lt}$, etc.), and let $H_l$ be the null
  hypothesis that $y_t - \hat{y}_{0t}$ is an \mds\ with respect to
  $F_{lt}$.  Also assume that the long-run variance-covariance matrix
  of the vector $(\bar{f}_1,\dots,\bar{f}_m)'$ is uniformly positive
  definite.

  Let the $*$-superscript denote a random variable with the
  distribution induced by the stationary bootstrap of
  Lemma~\ref{res:3}, letting $X_t = (y_{t+1}, x_t', \hat{y}_{1,t+1},
  \dots, \hat{y}_{m, t+1})'$ for $t > R$ (and unchanged for $t \leq
  R$).  Define $M_j$ as follows:
  \begin{enumerate}
  \item define $M_0 = \emptyset$,
  \item for $j = 1,2,\dots,m$, let $M_j = \{i : \bar{f}_{i} /
    \hat{\sigma}_i > \hat{d}_j\}$, where $\hat{d}_j$ is the $1-\alpha$
    quantile of $\max_{i \notin M_{j-1}} (\bar{f}_{i}^{*} -
    \bar{f_i}(\hat{\beta}_{T+1})) / \hat{\sigma}_i^{*}$,
  \end{enumerate}
  and reject all of the hypotheses $H_i$ with $i \in \bigcup_{j=0}^m
  M_j$.

  If $b \to \infty$ and $b^2/P \to 0$, this algorithm controls the
  \fwe\ at level $\alpha$ for the family of null hypotheses $H_i$
  ($i=1,\dots,m$): the probability of rejecting any true $H_i$ is less
  than or equal to $\alpha$.  Moreover, the probability that $H_i$ is
  rejected converges to one if $y_{t} - \hat{y}_{0t}(\beta_0)$ and
  $\hat{y}_{it}$ have nonzero correlation.
\end{thm}

@@

In this appendix, let $\to^{p^{*}}$ and $\to^{d^{*}}$ refer to
convergence in probability or distribution conditional on the observed
data.  Similarly, $\E^{*}$, $\var^{*}$, and $\cov^{*}$ refer to the
expectation, variance, and covariance with respect to the probability
measure induced by the stationary bootstrap, and $y_t^{*}$, etc. is
the random variable $y_t$ but under the bootstrap-induced \cdf.

\section{General Assumptions}

This section's lemmas depend on the following assumptions, which are
generalizations of the conditions of Theorems~\ref{res:1}
and~\ref{res:2}.  See \citet{Wes:96,Wes:06}, \citet{WeM:98}, and
\citet{Mcc:00} for a discussion of these conditions, as theirs are
nearly identical.

\begin{asmp}\label{asmp:a1}  We need two versions of the weak-dependence and moment
  conditions; the first for the asymptotic distribution of the \oos\
  statistic, and the second for the bootstrap distribution.
  \begin{enumerate}
  \item[a.] Let $\{y_t, z_t\}$ be an $L_2$-\ned\ process of size $-1/2$ on
    a strong mixing series $\{V_t\}$ of size $-r/(r-2)$, with $r > 2$.
  \item[b.] Let $\{y_t, z_t\}$ be an $L_{2+\delta}$-\ned\ process of
    size $-1$ on a strong mixing series $\{V_t\}$ of size $-(2r +
    \delta)/(r - 2)$, with $r > 2$ and $\delta > 0$.
  \end{enumerate}
\end{asmp}

@@

@@

\begin{lema}\label{res:a5}
  If Assumptions~\ref{asmp:a1}a, \ref{asmp:a2}a, \ref{asmp:a3},
  \ref{asmp:a4}, and \ref{asmp:a5} hold and $P \to \infty$ as $T \to
  \infty$, (add p.d. assumption too) then
  \begin{equation}
    P^{-1/2} \sum_{t=R+1}^T
    (f_{t}(\hat{\beta}_{t}) - \E f_{t}(\beta_{0})) \to^d N(0, \Omega),
  \end{equation}
  with
  \begin{align}
    \Omega =     
  \end{align}
\end{lema}
\begin{proof}[Proof of Lemma~\ref{res:a5}]
  Let $R'$ satisfy $R' \to \infty$ as $T \to \infty$ and $R'/T \to 0$
  and define $P' = T - R'$.  Then we can replace $P$ and $R$ with $P'$
  and $R'$ in the preceding equations without changing the asymptotic
  distributions of the statistics.  Note that $P'/R' \to \infty$.  The
  proof then follows exactly as in \citet[Theorem 2.3.1]{Mcc:00}
\end{proof}

The next Lemma presents our recursive window bootstrap scheme.
\begin{lema}\label{res:a6}
  If Assumptions~\ref{asmp:a1}b, \ref{asmp:a2}b, \ref{asmp:a3},
  \ref{asmp:a4}, and \ref{asmp:a5} hold , $P \to \infty$ as $T \to
  \infty$, and (p.d. assumption holds) then
  \begin{equation}
    P^{-1/2} \sum_{t=R+1}^T
    (f_{t}^{*}(\hat{\beta}_{t}^{*}) - \E^{*} f^{*}_{t}(\beta_{0}^{*})) \to^{d^{*}} N(0, \Omega^{*})
  \end{equation}
  and $\Omega^{*} \to^p \Omega$, where
  \begin{align}
    \Omega^{*} =     
  \end{align}
\end{lema}
